A Switching Planner for Combined Task and Observation Planning

### 2010-11-29 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/dtpddl.tex
OKay

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL was used in
all 3 of the International Planning Competitions since 2004. A
variation on PDDL for describing domains with stochastic actions and
uncertain starting configurations, PPDDL is a declarative first-order
language that facilitates factored description of domains and problem
instances. There are straightforward compilations from problems
expressed compactly in PPDDL to propositional representations amenable
to state-of-the-art planning procedures.

Because PPDDL cannot model domains that feature partial observability,
we developed an extension of that language, called DTPDDL. This can
express probabilistic models of the sensing consequences of acting, to
quantitatively capture unreliability in perception. That expressive
power is achieved by incorporating perceptual analogues of fluent,
predicate, and action definitions. In detail, we have declarations of
state characterising predicate and fluent symbols according to the
PPDDL syntax. In addition, we allow two other declarations, for
perceptual predicates and fluents respectively. For example, suppose
our robot is tasked with exploring {\em locations} in order to
identify the whereabouts of a {\em visual-object}. We must describe
state and perceptual facts that model the {\em true}, resp. perceived,
locations of objects. In DTPDDL, these declarations appear as:

@@

\small
\begin{tabtt}
(\=:action look-for-object \+ \\
   :parameters (\=?r - robot ?v - visual-object\\
   \> ?l - location) \\
   :precondition (and (= (is-in ?r) ?l) ) \\
   :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
\normalsize

\noindent In a slight departure from PPDDL, we suppose that
$\pp{look-for-object}$ has an effect on the state. That is, its
execution incurs an instantaneous penalty, $3$, that corresponds to
the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
reward} is a {\em reserved word}, and occurs in {\em increased} and
{\em decreased} terms in operator schemata. In that setting, {\em
reward} is interpreted as accumulated (resp. instantaneous) rewards
and penalties.}

Completing the example, we model the sensing outcome that results from
instantiating the $\pp{look-for-object}$ operator:

\small
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> look-for-object ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (o-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (o-is-in ?v) ?l))))) \\
\end{tabtt}
\normalsize

\noindent Here, the {\em execution} declaration is a lifted description of
the sense action-precondition -- I.e,
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office}) =
\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid us when discussing
our switching continual planner. The distribution is expressed in a
tree-like structure of terms. Each term is either: (1) atomic, e.g., a
state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$,
(2) probabilistic, e.g., $(\pp{probabilistic}~p_1 (T_1) .. p_n (T_n))$
where $T_i$ are conjunctive, or (3) a conjunct over probabilistic and
atomic terms. The root term is always conjunctive, and the leaves are
atomic. For example, the starting distribution for $\pp{DORA}$ might
be described thusly:\footnote{In PDDL, $(\pp{:init}~T_1..T_n)$
expresses the conjunctive root of the tree -- I.e., the root node
$(\pp{and}~T_1..T_n)$. Also, we shall write $\prop$, rather than
$(\pp{and}~\prop)$, it a conjunctive term contain a single atomic
subterm.}

@@

\noindent The interpretation of such an expression can be given
according to a {\em visitation} of terms. An atom is {\em visited} iff
its conjunctive parent is visited, and a conjunctive term is visited
iff all its immediate subterms are visited. A probabilistic term is
visited iff its conjunctive parent is visited, and exactly one of its
subterms, $T_i$, is visited. Each visitation of the root term
according to this recursive definition encapsulates a starting state,
along with the probability it occurs. The former corresponds to the
union of all visited atoms, and the latter corresponds to the product
of $p_i$ entries on the visited subterms of probabilistic
elements. Making these ideas concrete, our example yields the
following flat state distribution:

@@

>>>>

The modelling language of choice for planning in probabilistic
problems is the Probabilistic Planning Domain Definition
Language~\cite{younes:littman:04,younes:etal:2005}. PPDDL has been
used in all three of the International Planning Competitions since
2004. A variation on PDDL for describing domains with stochastic
actions and uncertain starting configurations, PPDDL is a declarative
first-order language that facilitates factored descriptions of domains
and problems. There are straightforward compilations from problems
expressed in PPDDL to propositional representations amenable
to state-of-the-art planners.

Because PPDDL cannot model domains that feature partial observability,
we develop an extension we call {\em Decision-Theoretic
(DT)PDDL}. This can express probabilistic models of the sensing
consequences of acting, to quantitatively capture unreliability in
perception. That expressive power is achieved by incorporating
perceptual analogues of fluent, predicate, and action definitions. In
detail, we have declarations of state characterising predicate and
fluent symbols according to the PPDDL syntax. In addition, we allow
two other declarations, for perceptual predicates and fluents
respectively. For example, suppose our robot is tasked with exploring
{\em locations} in order to identify the whereabouts of a {\em
visual-object}. We must describe state and perceptual facts that model
the {\em true}, resp. perceived, locations of objects. In DTPDDL,
these declarations appear as:

@@

\small
\begin{tabtt}
(\=:action object-search \+ \\
   :parameters (\=?r - robot ?v - visual-object\\
   \> ?l - location) \\
   :precondition (and (= (is-in ?r) ?l) ) \\
   :effect (and (assign (reward) -3) ) ) \\
\end{tabtt}
\normalsize

\noindent In a slight departure from PPDDL, we suppose that
$\pp{object-search}$ has an effect on the state. That is, its
execution incurs an instantaneous penalty, $3$, that corresponds to
the {\em cost} of performing the visual search.\footnote{In PPDDL {\em
reward} is a {\em reserved word}, and occurs in {\em increased} and
{\em decreased} terms in operator schemata. In that setting, reward is
accumulated, whereas in DTPDDL it is instantaneous.}

Completing the example, we model the sensing that results from
instantiating the $\pp{object-search}$ operator:

\small
\begin{tabtt}
(\= :sense vision \+\\
 :parameters \= (\= ?r - robot ?v - visual-object\\
 \>\>  ?l - location) \\
 :execution \> ( \> object-search ?r ?v ?l) \\
 :precondition (and (= (is-in ?r) ?l) ) \\
 :effect \>  (  \> and (when (= (is-in ?v) ?l) \\
   \> \> (probabilistic 0.8 \\
   \>  \>(= (o-is-in ?v) ?l))) \\
  \> (when (not (= (is-in ?v) ?l)) \\
   \>  \> (probabilistic 0.1 \\
   \>  \> (= (o-is-in ?v) ?l))))) \\
\end{tabtt}
\normalsize

\noindent Here, the {\em execution} declaration is a lifted description of
the sense action-precondition -- i.e,
$(\pp{object-search}~\pp{DORA}~\pp{box}~\pp{office})$ is equal to
$\poss_\stochActions(\pp{vision}~\pp{DORA}~\pp{box}~\pp{office})$. Above,
we include a redundant state-precondition, $(=(\pp{is-in}~?r)?l)$ in
order to fully demonstrate the syntax. Interpreting the above schema,
if action $(\pp{object-search}~\pp{DORA}~\pp{box}~\pp{office})$ is
executed, there is a $0.8$ chance of perceiving the box if it is in
the $\pp{office}$, and otherwise a $0.1$ chance of perceiving it.

The syntax and semantics for describing initial state distributions in
DTPDDL is taken verbatim from PPDDL. It is useful to present that
factored tree-like structure here, in order to aid us when discussing
our switching continual planner. The distribution is expressed in a
tree-like structure of terms. Each term is either: (1) atomic, e.g., a
state proposition such as $(=(\pp{is-in}~\pp{box})~\pp{office})$, (2)
probabilistic, e.g., $(\pp{probabilistic}~\prob_1 (T_1) .. \prob_n
(T_n))$ where $T_i$ are conjunctive, or (3) a conjunct over
probabilistic and atomic terms. The root term is always conjunctive,
and the leaves are atomic. For example, the starting distribution for
$\pp{DORA}$ might be described as follows:\footnote{In PDDL,
$(\pp{:init}~T_1..T_n)$ expresses the conjunctive root of the tree --
i.e., the root node $(\pp{and}~T_1..T_n)$. Also, we shall write
$\prop$, rather than $(\pp{and}~\prop)$, for conjunctive terms
that contain a single atomic subterm.}

@@

\noindent The interpretation of such an expression can be given
by a {\em visitation} of terms: An atom is {\em visited} iff
its conjunctive parent is visited, and a conjunctive term is visited
iff all its immediate subterms are visited. A probabilistic term is
visited iff its conjunctive parent is visited, and exactly one of its
subterms, $T_i$, is visited. Each visitation of the root term
according to this recursive definition encapsulates a starting state,
along with the probability that it occurs. The former corresponds to the
union of all visited atoms, and the latter corresponds to the product
of $\prob_i$ entries on the visited subterms of probabilistic
elements. Making these ideas concrete, our example yields the
following flat state distribution:

@@

From hereon, to simplify the discussion (and implementation)
we shall restrict our attention to POMDPs with deterministic
actions. We note that POMDPs with stochastic
actions can be compiled into equivalent deterministic-action POMDPs,
where all the original action uncertainty is expressed in the
starting-state distribution~\cite{ng:Jordan:2000}. In our setting,
that of finite-horizon planning, such a compilation yields a finite
flat representation of the original POMDP, only with deterministic
actions.
